### 2025-09-25 paper/sec/evaluation.tex
Better intro to 4.

We investigated the following research questions, which probe the limits of the \InterpretationAgent's ability
to reliably provide a data-driven interpretation of natural language. By contrasting genuine uncertainty with
misleading or ill-specified inputs, we aim to understand not only the model's raw success rate, but also the
failure modes most likely to arise in practical use.

@@

To evaluate the system we used a sample of the \SciGen dataset~\citep{moosavi21}. We aggregated the generated
expression in the linguistic categories from Table~\ref{tab:natural-language-forms}.
Figure~\ref{fig:success_rate_by_category} illustrates the success rate for each category, both with and
without target-value sharing. The results show that providing the target value (text fragment or numerical
value) is crucial for achieving a high success rate. Moreover, the \InterpretationAgent generates a correct
expression in approximately in 74.9\% of the cases (S.D. 3.0\%).

>>>>

Our evaluation tests the ability of the \InterpretationAgent to translate quantitative and semi-quantitative
expressions from scholarly natural language into executable queries that operate on the underlying dataset.
Beyond raw accuracy, we are also concerned with how performance varies with task complexity, and whether the
generated expressions are robust under changes to data or in the presence of ambiguity or other low data
quality issues. These are captured in two research questions:

@@

To evaluate the system we used a sample of the \SciGen dataset~\citep{moosavi21}. We aggregated the generated
expression in the linguistic categories from Table~\ref{tab:natural-language-forms}.
Figure~\ref{fig:success_rate_by_category} illustrates the success rate for each category, both with and
without target-value sharing. The results show that providing the target value (text fragment or numerical
value) is crucial for achieving a high success rate. Overall the \InterpretationAgent generates a correct
expression in approximately in 74.9\% of the cases (S.D. 3.0\%).
